\LARGE \textbf{Constrained Geometric Attitude Control on \( \SO \)}

### 2017-01-13 Response.tex

\end{correction}

In addition,~\Cref{prop:config_error} was modified to include additional detail on the notation \( \dirDiff{f}{x}\) as
\begin{correction}
The variation of \( A(R) \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}
		\dirDiff{A}{R} &= \eta \cdot e_{R_A} ,
	\end{align}
	where the notation \( \dirDiff{A}{R} \) represents the directional derivative of $A$ with respect to $R$ along $\delta R$.
	\end{correction}

\item \textit{This is serious. In~\cref{sec:attitude_control}, the upper bounds listed in~\cref{prop:eR_dot_bound} are given. Well, these should be the properties that your controller guarantees not come from the set D you've defined. Since these properties are used to derive the bound H which is used to derive the result in~\cref{prop:adaptive_control}, the stability proof under the proposed controller is far from complete.}

The stability proof has been improved and several modifications have been included in the manuscript:
\begin{itemize}
    \item An additional quadratic property has been derived for \( \Psi \) and shown in~\cref{prop:config_error},
    \item The quadratic property of \( \Psi \) requires the derivation of the hessian of \( \Psi \) given in~\cref{proof:config_error},
    \item Additional detail has been included on the set \( D \) and it's properties in~\cref{proof:config_error},
    \item The Lyapunov function in~\cref{proof:adaptive_control} is shown to be positive definite,
    \item The revised stability proof includes a modified requirement for the constant \( c \) shown in~\cref{eqn:c_bound}.
\end{itemize}

### 2017-01-16 manuscript_new.tex
Add differences between v1.0 and v2.0

\end{multicols}
\begin{figure}[t]
    \centering 
    \subfigure[{Configuration error \( \Psi \)} \label{fig:Psi_tv} ]{\includegraphics[width=0.3\columnwidth]{Psi_tv} }~
    \subfigure[{Angle to constraint } \label{fig:con_angle_tv}]{\includegraphics[width=0.3\columnwidth]{figures/constraint_angle_tv.pdf} }~
    \subfigure[{Disturbance estimate \(\bar\Delta\) components} \label{fig:Delta_tv}]{\includegraphics[width=0.3\columnwidth]{figures/delta_tv} }
    \caption{Time-varying external disturbance simulation}
    \label{fig:tv} 
\end{figure}
\begin{multicols}{2}
\subsection{Time-varying Disturbance}\label{ssec:time_varying}

@@

We demonstrate the ability to handle an uncertain time-varying disturbance via numerical example.
The system is identical to the one presented in~\Cref{sec:numerical_simulation}, however we modify the external disturbance. 
The external disturbance is the superposition of constant and time-varying terms as
\begin{align*}
    \Delta = \begin{bmatrix} 0.2 \\ 0.2 \\0.2 \end{bmatrix} + 0.02 \begin{bmatrix} \sin 9 t \\ \cos 9 t \\ \frac{1}{2} \parenth{\sin 9t + \cos 9t}\end{bmatrix} \si{\newton\meter}.
\end{align*}
We define a constraint in the inertial frame as \( v = [\frac{1}{\sqrt{2}}, \frac{1}{\sqrt{2}}, 0]^T \) with \( \theta = \ang{12} \).
The initial state is defined as \(R(0) = \exp( \frac{\pi}{2} \hat{e}_3) \), while the desired state is \(R_d =I \).
The goal is to rotate the vehicle about the \( e_3 \) axis while avoiding the obstacle and compensating for the time-varying disturbance.

@@

%To show~\cref{item:prop_crit} we note that \( A(R) \) is a positive definite function about \( R = R_d\), with \( \Psi(R_d) = A(R_d) = 0 \).
%As a result, there exists a neighborhood \(  U \) about \( R = R_d \) such that
%\begin{align*}
%  A(R) \geq A(R_d) \; \forall \; R \in U .
%\end{align*}
%Since the desired attitude, \( R_d \), must lie in the feasible set, i.e., \( r^T R^T v < \cos \theta \), the addition of the repulsive barrier function, \( B \), does not change the minimum value of \( \Psi \).
We show~\cref{item:prop_psi_quadratic} by computing the hessian of \( \Psi \), namely \( \Hess(\Psi) \),  using the chain rule as 
\begin{align*}
    \Hess &(\Psi)\cdot (\delta R,\delta R)= (\mathbf{D}_R\parenth{\mathbf{D}_R A \cdot \delta R}\cdot \delta R) B\\
    & + \parenth{\dirDiff{A}{R} }\parenth{\dirDiff{B}{R}} 
    + \parenth{\dirDiff{B}{R} }\parenth{\dirDiff{A}{R}}\\ &+(\mathbf{D}_R\parenth{\mathbf{D}_R B \cdot \delta R}\cdot \delta R) A.
\end{align*}
The first order derivative of \( A(R) \) and \( B(R) \) are given by~\cref{eqn:eRA,eqn:eRB}. 
The hessian of \( A(R) \) is computed as 
\begin{align*}
    \dirDiff{\parenth{\dirDiff{A}{R}}}{R} &= \eta \cdot \frac{1}{2}\parenth{G R_d^T \delta R - \delta R^T R_d G}^\vee , \\
    &= \eta \cdot \frac{1}{2} \parenth{G R_d^T R \hat \eta + \hat \eta R^T R_d G}^\vee,\\
    &= \eta \cdot \frac{1}{2} \bracket{\parenth{\braces{\tr{R^T R_d G} I - R^T R_d G} \eta}^\wedge}^\vee , \\
    &= \eta \cdot \frac{1}{2} \parenth{\tr{R^T R_d G} I - R^T R_d G}\eta ,
\end{align*}
where we used the scalar triple product rule, from~\cref{eqn:STP}, to arrive at the final form.
The hessian of \( B(R) \) is computed as
\begin{align*}
    \dirDiff{\parenth{\dirDiff{B}{R}}}{R} =& \eta \cdot \left[\frac{\parenth{\delta R^T v}^\wedge r}{\alpha\parenth{r^T R^T v - \cos \theta}} \right. \\
    & \left. - \frac{\parenth{R^T v}^\wedge r \parenth{r^T \delta R^T v}}{\alpha \parenth{r^TR^Tv - \cos \theta}^2}   \right] .
\end{align*}
The term \( \parenth{\delta R^T v}^\wedge r\) is simplified as
\begin{align*}
    \parenth{\delta R^T v}^\wedge r &= - \hat{r} \delta R ^T v , \\
    &= - \hat{r} \parenth{R \hat \eta}^T v , \\
    &= \hat{r} \parenth{\hat{\eta} R^T } v , \\
    &= - \hat{r} \parenth{R^T v}^\wedge \eta ,
\end{align*}
where we utilized the hat map property from~\cref{eqn:cross_product}.
Similarly, the term \( r^T \delta R^T v \) is simplified as
\begin{align*}
    r^T \delta R^T v &= r^T \parenth{-\hat \eta R^T}v , \\
    &= r^T \parenth{R^T v}^\wedge \eta .
\end{align*}
The hessian of \( B(R) \) then becomes
\begin{align*}
    \dirDiff{\parenth{\dirDiff{B}{R}}}{R} =& \eta \cdot \left[ - \frac{ \hat r \parenth{R^T v}^\wedge}{\alpha \parenth{r^T R^T v - \cos \theta}} \right. \\ 
    & \left. - \frac{\parenth{R^T v}^\wedge r r^T \parenth{R^T v}^\wedge }{ \alpha\parenth{r^T R^T v - \cos \theta}^2}  \right] \eta .
\end{align*}
Using these terms, we evaluate \( \Hess{\Psi} \) at the desired attitude \( R = R_d \) as follows. Since $A=0$ and $\mathbf{D}_R A=0$ at $R=R_d$, 
\begin{align*}
    \Hess(\Psi)\cdot (\delta R,\delta R)\big|_{R=R_d} = \eta \cdot \frac{1}{2} B \parenth{\tr{G}I -  G} \eta , 
\end{align*}
which is positive definite since \( B > 1\) and \( \sum g_i > g_i\). 
The domain \( D \) is an open neighborhood of the desired attitude \( R_d \), and it excludes the undesired equilibrium points of \( A(R) \) and the infeasible regions defined by the constraints \( r^T R^T v_i \). 
Therefore, the only critical point of the error function $\Psi$ in the domain $D$ correspond to the desired attitude $R=R_d$ with $e_R=0$ and $\Psi=0$. 
Therefore, in \( D \) the configuration error function is quadratic and the bounds in~\cref{item:prop_psi_quadratic} are valid according to~\cite[Proposition 6.30]{bullo2004}.

>>>>

\end{multicols}
\begin{figure}[h]
    \centering 
    \subfigure[{Configuration error \( \Psi \)} \label{fig:Psi_tv} ]{\includegraphics[width=0.3\columnwidth]{Psi_tv} }~
    \subfigure[{Angle to constraint } \label{fig:con_angle_tv}]{\includegraphics[width=0.3\columnwidth]{figures/constraint_angle_tv.pdf} }~
    \subfigure[{Disturbance estimate \(\bar\Delta\) components} \label{fig:Delta_tv}]{\includegraphics[width=0.3\columnwidth]{figures/delta_tv} }
    \caption{Time-varying external disturbance simulation}
    \label{fig:tv} 
\end{figure}
\begin{multicols}{2}
\subsection{Time-varying Disturbance}\label{ssec:time_varying}

@@

We demonstrate the ability to handle an uncertain time-varying disturbance via numerical example.
The system is identical to the one presented in~\cref{sec:numerical_simulation}, however we modify the external disturbance. 
The external disturbance is the superposition of constant and time-varying terms as
\begin{align*}
    \Delta = \begin{bmatrix} 0.2 \\ 0.2 \\0.2 \end{bmatrix} + 0.02 \begin{bmatrix} \sin 9 t \\ \cos 9 t \\ \frac{1}{2} \parenth{\sin 9t + \cos 9t}\end{bmatrix} \si{\newton\meter}.
\end{align*}
We define a constraint in the inertial frame as \( v = [\frac{1}{\sqrt{2}}, \frac{1}{\sqrt{2}}, 0]^T \) with \( \theta = \ang{12} \).
The initial state is defined as \(R(0) = \exp( \frac{\pi}{2} \hat{e}_3) \), while the desired state is \(R_d =I \).
The goal is to rotate the vehicle about the \( e_3 \) axis while avoiding the obstacle and compensating for the time-varying disturbance.

@@

%To show~\cref{item:prop_crit} we note that \( A(R) \) is a positive definite function about \( R = R_d\), with \( \Psi(R_d) = A(R_d) = 0 \).
%As a result, there exists a neighborhood \(  U \) about \( R = R_d \) such that
%\begin{align*}
%  A(R) \geq A(R_d) \; \forall \; R \in U .
%\end{align*}
%Since the desired attitude, \( R_d \), must lie in the feasible set, i.e., \( r^T R^T v < \cos \theta \), the addition of the repulsive barrier function, \( B \), does not change the minimum value of \( \Psi \).
We show~\cref{item:prop_psi_quadratic} by computing the hessian of \( \Psi \), namely \( \Hess(\Psi) \),  using the chain rule as 
\begin{align*}
    \Hess &(\Psi)\cdot (\delta R,\delta R)= (\mathbf{D}_R\parenth{\mathbf{D}_R A \cdot \delta R}\cdot \delta R) B\\
    & + \parenth{\dirDiff{A}{R} }\parenth{\dirDiff{B}{R}} 
    + \parenth{\dirDiff{B}{R} }\parenth{\dirDiff{A}{R}}\\ &+(\mathbf{D}_R\parenth{\mathbf{D}_R B \cdot \delta R}\cdot \delta R) A.
\end{align*}
The first order derivative of \( A(R) \) and \( B(R) \) are given by~\cref{eqn:eRA,eqn:eRB}. 
The hessian of \( A(R) \) is computed as 
\begin{align*}
    \dirDiff{\parenth{\dirDiff{A}{R}}}{R} &= \eta \cdot \frac{1}{2}\parenth{G R_d^T \delta R - \delta R^T R_d G}^\vee , \\
    &= \eta \cdot \frac{1}{2} \parenth{G R_d^T R \hat \eta + \hat \eta R^T R_d G}^\vee,\\
    &= \eta \cdot \frac{1}{2} \bracket{\parenth{\braces{\tr{R^T R_d G} I - R^T R_d G} \eta}^\wedge}^\vee , \\
    &= \eta \cdot \frac{1}{2} \parenth{\tr{R^T R_d G} I - R^T R_d G}\eta ,
\end{align*}
where we used the scalar triple product rule, from~\cref{eqn:STP}, to arrive at the final form.
The hessian of \( B(R) \) is computed as
\begin{align*}
    \dirDiff{\parenth{\dirDiff{B}{R}}}{R} =& \eta \cdot \left[\frac{\parenth{\delta R^T v}^\wedge r}{\alpha\parenth{r^T R^T v - \cos \theta}} \right. \\
    & \left. - \frac{\parenth{R^T v}^\wedge r \parenth{r^T \delta R^T v}}{\alpha \parenth{r^TR^Tv - \cos \theta}^2}   \right] .
\end{align*}
The term \( \parenth{\delta R^T v}^\wedge r\) is simplified as
\begin{align*}
    \parenth{\delta R^T v}^\wedge r &= - \hat{r} \delta R ^T v , \\
    &= - \hat{r} \parenth{R \hat \eta}^T v , \\
    &= \hat{r} \parenth{\hat{\eta} R^T } v , \\
    &= - \hat{r} \parenth{R^T v}^\wedge \eta ,
\end{align*}
where we utilized the hat map property from~\cref{eqn:cross_product}.
Similarly, the term \( r^T \delta R^T v \) is simplified as
\begin{align*}
    r^T \delta R^T v &= r^T \parenth{-\hat \eta R^T}v , \\
    &= r^T \parenth{R^T v}^\wedge \eta .
\end{align*}
The hessian of \( B(R) \) then becomes
\begin{align*}
    \dirDiff{\parenth{\dirDiff{B}{R}}}{R} =& \eta \cdot \left[ - \frac{ \hat r \parenth{R^T v}^\wedge}{\alpha \parenth{r^T R^T v - \cos \theta}} \right. \\ 
    & \left. - \frac{\parenth{R^T v}^\wedge r r^T \parenth{R^T v}^\wedge }{ \alpha\parenth{r^T R^T v - \cos \theta}^2}  \right] \eta .
\end{align*}
Using these terms, we evaluate \( \Hess{\Psi} \) at the desired attitude \( R = R_d \) as follows. Since $A=0$ and $\mathbf{D}_R A=0$ at $R=R_d$, 
\begin{align*}
    \Hess(\Psi)\cdot (\delta R,\delta R)\big|_{R=R_d} = \eta \cdot \frac{1}{2} B \parenth{\tr{G}I -  G} \eta , 
\end{align*}
which is positive definite since \( B > 1\) and \( \sum g_i > g_i\). 
The domain \( D \) is an open neighborhood of the desired attitude \( R_d \), and it excludes the undesired equilibrium points of \( A(R) \) and the infeasible regions defined by the constraints \( r^T R^T v_i \). 
Therefore, the only critical point of the error function $\Psi$ in the domain $D$ corresponds to the desired attitude $R=R_d$ with $e_R=0$ and $\Psi=0$. 
Therefore, in \( D \) the configuration error function is quadratic and the bounds in~\cref{item:prop_psi_quadratic} are valid according to~\cite[Proposition 6.30]{bullo2004}.
